---
title: '`r report_filename`'
author: |
  | `r report_author`
  | ENPRC Genomics Core"
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
    fig_width: 9
    toc_depth: 3
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 8, fig.height = 6) 

library(kableExtra)   ## Needed if rendering tables
```

<!-- This file will not execute by itself. It should be called within the  -->
<!-- runfile to render the final report. You can arrange figures, add explanations, -->
<!-- and do other aesthetic tweaks here. -->

```{=html}
<style type="text/css">
.book .book-body .page-inner {
  max-width: 1600px;
  margin-left: auto;
  margin-right: auto;
}
details > summary {
  display: list-item;
}
</style>
```

In this report, we review some of the key quality control metrics calculated by cellranger.

# Gene expression

```{r, fig.width=14, fig.height=6}
qc_metrics_gex %>%
  ggplot(aes(x=ID, y=metric_value, fill=ID)) +
    geom_col() +
    facet_wrap(~metric_name,
               scales='free_y',
               ncol=2,
               labeller = labeller(metric_name = label_wrap_gen(100))) +
    theme_bw() +
    labs(x = 'Sample', y = 'Value', title = 'GEX') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

```{r}
qc_metrics_gex %>%
  pivot_wider(names_from = metric_name, values_from = metric_value) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(full_width = F)
```

# Antibody-derived tags

[Explanation of metrics](https://support.10xgenomics.com/single-cell-gene-expression/software/pipelines/latest/output/antibody-metrics)

<details open><summary>**Percent Antibody Reads Usable:**</summary>

Reads that contain a recognized antibody barcode, a valid UMI, and a cell-associated barcode.

Ideal > 20%. This may indicate poor library quality for the antibody library, poor sequencing quality, or mistakes while specifying antibody details in the Feature Reference CSV provided to Cell Ranger.

</details>

```{r, fig.width=14, fig.height=6}
qc_metrics_adt %>%
  ggplot(aes(x=ID, y=metric_value, fill=ID)) +
    geom_col() +
    facet_wrap(~metric_name,
               ## you may prefer fixed Y scales depending on what you want to show
               scales='free_y',
               ncol=2,
               labeller = as_labeller(
                  ## option to rename some metrics to add clarity
                 c(`Cells` = 'Cells',
                   `Fraction antibody reads usable` = 'Percent antibody reads usable',
                   `Mean antibody reads usable per cell` = 'Mean antibody reads usable per cell',
                   `Median genes per cell` = 'Median genes per cell'))) +
    theme_bw() +
    labs(x = 'Sample', y = 'Value', title = 'Antibody Capture') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

```{r}
qc_metrics_adt %>%
  pivot_wider(names_from = metric_name, values_from = metric_value) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(full_width = F)
```

<!-- *** -->

<!-- # VDJ B libraries -->

<!-- ## VDJ B -->

<!-- [Explanation of metrics](https://support.10xgenomics.com/single-cell-vdj/software/pipelines/latest/output/metrics) -->

<!-- <details><summary>**Number of Cells With Productive V-J Spanning Pair**</summary> -->

<!-- [Productive Contigs](https://support.10xgenomics.com/single-cell-vdj/software/pipelines/latest/algorithms/annotation#productive) -->

<!-- A contig is termed productive if the following conditions are met: -->

<!-- * **Full length requirement:** The contig matches the initial part of a V gene. The contig continues on, ultimately matching the terminal part of a J gene. -->

<!-- * **Start requirement:** The initial part of the V matches a start codon on the contig. Note that in the human and mouse reference sequences supplied by 10x Genomics, every V segment begins with a start codon. -->

<!-- * **Nonstop requirement:** There is no stop codon between the V start and the J stop. -->

<!-- * **In-frame requirement:** The J stop minus the V start equals one mod three. This just says that the codons on the V and J segments are in frame. -->

<!-- * **CDR3 requirement:** There is an annotated CDR3 sequence. -->

<!-- * **Structure requirement:** Let VJ denote the sum of the lengths of the V and J segments. Let len denote the J stop minus the V start, measured on the contig. Then VJ - len lies between -25 and +25, except for IGH, which must be between -55 and +25. This condition is imposed to preclude anomalous structure changes that are unlikely to correspond to functional proteins. -->
<!-- </details> -->

<!-- <details><summary>**Paired clonotype diversity**</summary> -->

<!-- Effective diversity of the paired clonotypes, computed as the Inverse Simpson Index of the clonotype frequencies. A value of 1 indicates a minimally diverse sample - only one distinct clonotype was detected. A value equal to the estimated number of cells indicates a maximally diverse sample. -->

<!-- </details> -->
