---
title: 'scRNA QC template'
author: "ENPRC Genomics Core"
date: '`r Sys.Date()`'
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.width = 8, fig.height = 6,
                      cache=FALSE, cache.lazy = FALSE) 

library(tidyverse)  ## General R logic
library(here)       ## Easier specification of file locations
library(yaml)       ## parses config yaml
```

This script performs quality checks on cellranger processed data. It is
not intended to perform any data manipulation. This should be lightweight,
but the format file is separated from the runfile for ease of use and 
manipulating report aesthetics. The runfile (current file) should be ran
all the way through, with the last step being to render the format file
based on objects in the environment produced by the runfile. 

This runfile is in a html_notebook format so you can view all the outputs
in a single HTML if that's useful, but the final report should be 
the one rendered by the format file into the reports folder (see final chunk).

You may have to adjust some filter/select functions here, as cellranger seems
to change their output naming conventions frequently. 

# Read config files
```{r}
config <- yaml.load_file(here("config/scRNA_GEX_config.yml"))
samplesheet <- read.table(here('config/samplesheet.txt'), header = TRUE)
```

The output formats from cellranger multi differ from cellranger count. Choose
the appropriate code based on your inputs.

# Cellranger Count

## ADT libraries

### Plot

```{r, fig.width=10}
## plots are saved to environment so they can be rendered in the format file
qc_plot_ADT <- qc_metrics %>%
  select(ID, contains('Antibody')) %>%
  ## Specify metrics of interest below
  select(ID,
         Antibody..Median.UMIs.per.Cell..summed.over.all.recognized.antibody.barcodes.,
         Antibody..Sequencing.Saturation,
         Antibody..Fraction.Antibody.Reads.Usable) %>%
  rename_with(~ gsub("Antibody..", "", .x, fixed=TRUE)) %>%
  rename_with(~ gsub(".", " ", .x, fixed=TRUE)) %>%
  reshape2::melt(id.vars = 'ID') %>%
  ggplot(aes(x=ID, y=value, fill=ID)) +
    geom_col() + 
    facet_wrap(~variable,
               ## you may prefer fixed Y scales depending on what you want to show
               scales='free_y',
               ncol=3,
               labeller = as_labeller(
                  ## option to rename some metrics to add clarity
                 c(`Median UMIs per Cell  summed over all recognized antibody barcodes ` = 'Median UMIs per cell',
                   `Sequencing Saturation` = 'Antibody sequencing saturation',
                   `Fraction Antibody Reads Usable` = 'Fraction antibody reads usable'))) +
    theme_bw() +
    labs(x = 'Sample', y = 'Value', title = 'Antibody Capture') +
  theme(axis.text.x = element_text(angle=45, hjust = 1))
qc_plot_ADT
```

### Table

```{r}
qc_table_ADT <- qc_metrics %>% 
 select(ID, contains('Antibody')) %>%
  ## Specify metrics of interest below
  select(ID,
         Antibody..Median.UMIs.per.Cell..summed.over.all.recognized.antibody.barcodes.,
         Antibody..Sequencing.Saturation,
         Antibody..Fraction.Antibody.Reads.Usable) %>%
  rename_with(~ gsub("Antibody..", "", .x, fixed=TRUE)) %>%
  rename_with(~ gsub(".", " ", .x, fixed=TRUE))
qc_table_ADT
# knitr::kable(qc_table_ADT) %>%
  # kable_styling(full_width = F)
```

## GEX libraries

### Plot

```{r, fig.width=14, fig.height=6}
qc_plot_GEX <- qc_metrics %>%
  select(ID, !contains('Antibody')) %>%
  rename_with(~ gsub(".", " ", .x, fixed=TRUE)) %>%
  # filter(Category == 'Cells') %>%
  select(ID,
         'Estimated Number of Cells',
         'Median Genes per Cell',
         'Median UMI Counts per Cell',
         'Reads Mapped Confidently to Genome',
         'Total Genes Detected',
         'Sequencing Saturation') %>%
  reshape2::melt(id.vars = 'ID') %>%
  ggplot(aes(x=ID, y=value, fill=ID)) +
    geom_col() + 
    facet_wrap(~variable, 
               scales='free_y',
               ncol=3
               #labeller = labeller(Metric.Name = label_wrap_gen(100))
               ) +
    theme_bw() +
    labs(x = 'Sample', y = 'Value', title = 'GEX') +
  theme(axis.text.x = element_text(angle=45, hjust = 1))
qc_plot_GEX
```

### Table

```{r}
qc_table_GEX <- qc_metrics %>% 
  select(ID, !contains('Antibody')) %>%
  rename_with(~ gsub(".", " ", .x, fixed=TRUE)) %>%
  # filter(Category == 'Cells') %>%
  select(ID,
         'Estimated Number of Cells',
         'Median Genes per Cell',
         'Median UMI Counts per Cell',
         'Reads Mapped Confidently to Genome',
         'Total Genes Detected',
         'Sequencing Saturation')
qc_table_GEX
```

## VDJ

I haven't encountered VDJ data from cellranger count yet, so don't have template code for that.
You could check the multi section for ideas

# Cellranger Multi

## Gather metric files

```{r}
## This may be moved to package eventually if it seems robust enough
capture_cellranger_multi_qc_metrics <- function(samplesheet, config){
  files <- lapply(samplesheet$FileID,
                  function(x) {read.csv(Sys.glob(file.path(
                    config$rootDir, config$alignmentDir,
                    x,"outs/per_sample_outs/",
                    x,"metrics_summary.csv")))})
  qc <- data.table::rbindlist(files, use.names=TRUE, fill=TRUE,idcol="ID")
  qc$ID <- plyr::mapvalues(qc$ID, from=1:length(samplesheet$Label), to=samplesheet$Label)
  qc
}

## Use the FileIDs from the samplesheet to find metric files in the
## cellranger output directory as specified in the config
qc_metrics <- capture_cellranger_multi_qc_metrics(samplesheet, config) 

## Some additional formatting is applied to complex values
qc_metrics <- qc_metrics%>%
  mutate(Metric.Value = readr::parse_number(Metric.Value)) %>%
  mutate(ID = factor(ID, levels = unique(gtools::mixedsort(ID)))) %>%
  arrange(ID)
```

Depending on the libraries in your captures, these may not all apply

## ADT libraries

### Plot

```{r, fig.width=10}
## plots are saved to environment so they can be rendered in the format file
qc_plot_ADT <- qc_metrics %>%
  filter(Library.Type == 'Antibody Capture') %>%
  ## Specify metrics of interest below
  filter(Metric.Name %in% c('Cells',
                            'Median UMI counts per cell',
                            'Fraction antibody reads usable',
                            'Mean antibody reads usable per cell')) %>%
  ggplot(aes(x=ID, y=Metric.Value, fill=ID)) +
    geom_col() + 
    facet_wrap(~Metric.Name,
               ## you may prefer fixed Y scales depending on what you want to show
               scales='free_y',
               ncol=3,
               labeller = as_labeller(
                  ## option to rename some metrics to add clarity
                 c(`Cells` = 'Cells',
                   `Fraction antibody reads usable` = 'Percent antibody reads usable',
                   `Mean antibody reads usable per cell` = 'Mean antibody reads usable per cell',
                   `Median UMI counts per cell` = 'Median UMI counts per cell'))) +
    theme_bw() +
    labs(x = 'Sample', y = 'Value', title = 'Antibody Capture') +
  theme(axis.text.x = element_text(angle=45, hjust = 1))
qc_plot_ADT
```

### Table

```{r}
qc_table_ADT <- qc_metrics %>% 
  filter(Library.Type=='Antibody Capture') %>% 
  filter(Metric.Name %in% c('Cells',
                            'Median UMI counts per cell',
                            'Fraction antibody reads usable',
                            'Mean antibody reads usable per cell')) %>%
  select(ID, Metric.Name, Metric.Value) %>% 
  reshape(idvar="ID",timevar = "Metric.Name", direction='wide') %>%
  rename_with(~ gsub("Metric.Value.", "", .x, fixed=TRUE))
qc_table_ADT
# knitr::kable(qc_table_ADT) %>%
  # kable_styling(full_width = F)
```

## GEX libraries

### Plot

```{r, fig.width=14, fig.height=6}
qc_plot_GEX <- qc_metrics %>%
  filter(Library.Type == 'Gene Expression') %>%
  # filter(Category == 'Cells') %>%
  filter(Metric.Name %in% c('Cells',
                            'Median genes per cell',
                            'Median UMI counts per cell',
                            'Confidently mapped reads in cells')) %>%
  ggplot(aes(x=ID, y=Metric.Value, fill=ID)) +
    geom_col() + 
    facet_wrap(~Metric.Name, 
               scales='free_y',
               ncol=3,
               labeller = labeller(Metric.Name = label_wrap_gen(100))) +
    theme_bw() +
    labs(x = 'Sample', y = 'Value', title = 'GEX') +
  theme(axis.text.x = element_text(angle=45, hjust = 1))
qc_plot_GEX
```

### Table

```{r}
qc_table_GEX <- qc_metrics %>% 
  filter(Library.Type=='Gene Expression') %>% 
  select(ID, Metric.Name, Metric.Value) %>%
  filter(Metric.Name %in% c('Cells',
                            'Median genes per cell',
                            'Median UMI counts per cell',
                            'Confidently mapped reads in cells')) %>%
  reshape(idvar="ID",timevar = "Metric.Name", direction='wide') %>%
  rename_with(~ gsub("Metric.Value.", "", .x, fixed=TRUE))
qc_table_GEX
```

## VDJ B libraries

VDJ data is not present in the example dataset, so these won't work, but
you get the idea. Set eval=TRUE if you do have vdj data

### Plot

```{r, fig.width=18, fig.height=4, eval=FALSE}
qc_plot_VDJB <- qc_metrics %>%
  filter(Library.Type=='VDJ B') %>%
  # filter(Category == 'Cells') %>%
  filter(Metric.Name %in% c('Estimated number of cells',
                            'Cells with productive V-J spanning pair',
                            'Paired clonotype diversity')) %>%
  ggplot(aes(x=ID, 
             y=Metric.Value, 
             fill=ID)) +
    geom_col() + 
    facet_wrap(~ factor(Metric.Name,
                        levels = c('Estimated number of cells',
                                 'Cells with productive V-J spanning pair',
                                 'Paired clonotype diversity')),
               scales='free_y',
               ncol=3,
               labeller = as_labeller(c(
                 `Estimated number of cells` = 'Estimated number of cells',
                 `Cells with productive V-J spanning pair` = 'Percent cells with productive V-J spanning pair',
                 `Paired clonotype diversity` = 'Paired clonotype diversity'))
               ) +
    theme_bw() +
    labs(x = 'Sample', y = 'Value', title = 'VDJ B')
qc_plot_VDJB
```

### Table

```{r, eval=FALSE}
qc_table_VDJB <- qc_metrics %>% 
  filter(Library.Type=='VDJ B') %>%
  # filter(Category == 'Cells') %>%
  filter(Metric.Name %in% c('Estimated number of cells',
                            'Cells with productive V-J spanning pair',
                            'Paired clonotype diversity')) %>%
  select(ID, Metric.Name, Metric.Value) %>%
  filter(Metric.Name %in% c('Cells',
                            'Median genes per cell',
                            'Median UMI counts per cell',
                            'Confidently mapped reads in cells')) %>%
  reshape(idvar="ID",timevar = "Metric.Name", direction='wide') %>%
  rename_with(~ gsub("Metric.Value.", "", .x, fixed=TRUE))
qc_table_VDJB
```

# Render report

```{r}
## This only works if all the variables called in the format file 
## are in the current environment. Check the format file to tweak
## aesthetics 
rmarkdown::render(here('analysis_scripts/01_QC-template.format.Rmd'),
                  output_file = '01_QC-report.html',
                  output_dir = here('reports'))
```
