---
title: "`r report_filename`"
author: |
  | `r report_author`
  | ENPRC Genomics Core
date: "`r Sys.Date()`"
output: 
  rmdformats::robobook:
    fig_width: 9
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=TRUE, message=TRUE,
                      fig.width = 8, fig.height = 6,
                      cache=FALSE, cache.lazy = FALSE,
                      fig.align = 'center')#, include=FALSE)
library(kableExtra)
```

```{=html}
<style type="text/css">
.book .book-body .page-inner {
  max-width: 1600px;
  margin-left: auto;
  margin-right: auto;
}
details > summary {
  display: list-item;
}
</style>
```

# Full dataset

Subsetting just APCs based on the labels from Report 03 
```{r, fig.width = 28, fig.height = 12}
ggarrange(
  DimPlot(s.full, reduction = "wnn.umap.harmony", group.by = "ADT_and_BP_labels.wnn.res1.low", 
        label = T, label.box = T, label.size = 4, repel = T) + 
    NoLegend() + theme(aspect.ratio = 1),
  DimPlot(s.full, reduction = "wnn.umap.harmony", group.by = "ADT_and_BP_labels.wnn.res1.high", 
        label = T, label.box = T, label.size = 4, repel = T) + 
    NoLegend() + theme(aspect.ratio = 1),
  DimPlot(s.full, reduction = "wnn.umap.harmony", group.by = "APC_subset", 
        label = T, label.box = T, label.size = 4, repel = T) + 
    NoLegend() + theme(aspect.ratio = 1),
  ncol=3, nrow=1
  )
```


# Integration

## Before integration {.tabset}

### AnimalID
```{r, fig.width = 12, fig.height = 6}
ggarrange(
  DimPlot(s.APC, reduction = "umap_rna", group.by = "animalID") + 
    theme(aspect.ratio=1) + ggtitle("RNA before Integration"),
  DimPlot(s.APC, reduction = "umap_adt_pca", group.by = "animalID") + 
    theme(aspect.ratio=1) + ggtitle("ADT before Integration"),
  nrow=1, ncol=2, common.legend = T, legend = "bottom"
)
```

### Group
```{r, fig.width = 12, fig.height = 6}
ggarrange(
  DimPlot(s.APC, reduction = "umap_rna", group.by = "group") + 
    theme(aspect.ratio=1) + ggtitle("RNA before Integration"),
  DimPlot(s.APC, reduction = "umap_adt_pca", group.by = "group") + 
    theme(aspect.ratio=1) + ggtitle("ADT before Integration"),
  nrow=1, ncol=2, common.legend = T, legend = "bottom"
)
```

### Challenge
```{r, fig.width = 12, fig.height = 6}
ggarrange(
  DimPlot(s.APC, reduction = "umap_rna", group.by = "challenge") + 
    theme(aspect.ratio=1) + ggtitle("RNA before Integration"),
  DimPlot(s.APC, reduction = "umap_adt_pca", group.by = "challenge") + 
    theme(aspect.ratio=1) + ggtitle("ADT before Integration"),
  nrow=1, ncol=2, common.legend = T, legend = "bottom"
)
```

### Report 03 labels
```{r, fig.width = 12, fig.height = 6}
ggarrange(
  DimPlot(s.APC, reduction = "umap_rna", group.by = "ADT_and_BP_labels.wnn.res1.low",
          label = T, label.box = T, label.size = 2, repel = T) + 
    theme(aspect.ratio=1) + ggtitle("RNA before Integration"),
  DimPlot(s.APC, reduction = "umap_adt_pca", group.by = "ADT_and_BP_labels.wnn.res1.low",
          label = T, label.box = T, label.size = 2, repel = T) + 
    theme(aspect.ratio=1) + ggtitle("ADT before Integration"),
  nrow=1, ncol=2, common.legend = T, legend = "bottom"
)
```

## Integration of either RNA or ADT {.tabset}

### AnimalID
```{r, fig.width = 12, fig.height = 6}
ggarrange(
  DimPlot(s.APC, reduction = "umap_harmony_rna", group.by = "animalID") + theme(aspect.ratio=1) + ggtitle("RNA Harmonized"),
  DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "animalID") + theme(aspect.ratio=1) + ggtitle("ADT Harmonized"),
  DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "animalID") + theme(aspect.ratio=1) + ggtitle("Both Harmonized (WNN)"),
  nrow=1, ncol=3, common.legend = T, legend = "bottom"
)
```

### Group
```{r, fig.width = 12, fig.height = 6}
ggarrange(
  DimPlot(s.APC, reduction = "umap_harmony_rna", group.by = "group") + theme(aspect.ratio=1) + ggtitle("RNA Harmonized"),
  DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "group") + theme(aspect.ratio=1) + ggtitle("ADT Harmonized"),
  DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "group") + theme(aspect.ratio=1) + ggtitle("Both Harmonized (WNN)"),
  nrow=1, ncol=3, common.legend = T, legend = "bottom"
)
```

### Challenge
```{r, fig.width = 12, fig.height = 6}
ggarrange(
  DimPlot(s.APC, reduction = "umap_harmony_rna", group.by = "challenge") + theme(aspect.ratio=1) + ggtitle("RNA Harmonized"),
  DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "challenge") + theme(aspect.ratio=1) + ggtitle("ADT Harmonized"),
  DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "challenge") + theme(aspect.ratio=1) + ggtitle("Both Harmonized (WNN)"),
  nrow=1, ncol=3, common.legend = T, legend = "bottom"
)
```

### Report 03 labels
```{r, fig.width = 12, fig.height = 6}
ggarrange(
  DimPlot(s.APC, reduction = "umap_harmony_rna", group.by = "ADT_and_BP_labels.wnn.res1.low",
          label = T, label.box = T, label.size = 2, repel = T) + 
    theme(aspect.ratio=1) + ggtitle("RNA Harmonized"),
  DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "ADT_and_BP_labels.wnn.res1.low",
          label = T, label.box = T, label.size = 2, repel = T) + 
    theme(aspect.ratio=1) + ggtitle("ADT Harmonized"),
  DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "ADT_and_BP_labels.wnn.res1.low",
          label = T, label.box = T, label.size = 2, repel = T) + 
    theme(aspect.ratio=1) + ggtitle("Both Harmonized (WNN)"),
  nrow=1, ncol=3, common.legend = T, legend = "bottom"
)
```


# Selecting a cluster resolution

## UMAPs {.tabset}

Below are the clusters calculated at different resolutions for the Weighted-Nearest Neighbors graph. The rest of the report will use resolution 0.3, but let me know if you'd like to change it.

```{r, results='asis'}
chunk_text_res_umaps <- c(
  '### {{res}}\n',
   '```{r, fig.width = 12, fig.height = 6}
   ggarrange(
    DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.{{res}}", label = T) + theme(aspect.ratio = 1),
    DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "ADT_and_BP_labels.wnn.res1.low", label = T, label.box = T, label.size = 2, repel = T) + theme(aspect.ratio = 1) + NoLegend(),
    ncol = 2, nrow = 1
   )',
   '```\n'
)

out <- NULL
for (res in seq(0.1, 1, 0.1)) {
  out <- c(out, knitr::knit_expand(text = chunk_text_res_umaps))
}
cat(knitr::knit_child(text = out, quiet = TRUE), sep = "\n")
```

## Cluster tree

The size of each node is related to the number of cells in each cluster and the color indicates the clustering resolution. Edges are colored according to the number of cells they represent and the transparency shows the incoming node proportion, the number of cells in the edge divided by the number of cells in the node it points to.

```{r, fig.width = 7, fig.height = 8}
clustree::clustree(s.APC, prefix = "wsnn_harmony_res.")
```

## Unsupervised clusters and ADT panel

### Surface protein marker definitions

```{r}
adt_short_name_tbl %>%
  select(short_name, marker_of_short) %>%
  knitr::kable() %>% kable_styling()
```

### Unsupervised clusters and ADT panel

```{r, fig.width = 16, fig.height = 10}
gridExtra::grid.arrange(
  DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.0.3",
          label = T, label.box = T, label.size = 6, repel = T) + 
    NoLegend() + theme(aspect.ratio=1) + ggtitle("Unsupervised clusters"),
  DotPlot(s.APC, features = adt_short_name_tbl$original_name, group.by = "wsnn_harmony_res.0.3", cluster.idents = T) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    scale_x_discrete(labels = paste0(adt_short_name_tbl$marker_of_short, " [", adt_short_name_tbl$short_name, "]")),
  widths = c(1.5,1))
```



# Unsupervised cluster marker heatmaps and lists {.tabset}

## RNA

Showing only the top 10 markers based on gene expression, ranked by p-value and then average log fold change.

### Heatmap of top 10 markers for each cluster

```{r, fig.width=9, fig.height=24}
Idents(s.APC) <- "wsnn_harmony_res.0.3"

markers$rna %>% 
  filter(p_val_adj < 0.0001, avg_log2FC > 0.25) %>% 
  arrange(cluster, p_val_adj, desc(avg_log2FC)) %>%
  group_by(cluster) %>% slice_head(n=10) %>% pull(gene) %>% unique() %>%
DoHeatmap(subset(s.APC, downsample=500), group.by = "wsnn_harmony_res.0.3", assay = "RNA", features = ., size = 4) + 
  theme(axis.text.y = element_text(size=8))
```

### Table of top 10 markers for each cluster

```{r}
markers$rna %>% 
  filter(p_val_adj < 0.0001, avg_log2FC > 0.25) %>% 
  arrange(cluster, p_val_adj, desc(avg_log2FC)) %>%
  group_by(cluster) %>% slice_head(n=10) %>%
  DT::datatable(extensions = "Buttons",
                options = list(dom = "Blfrtip",
                               buttons = c("copy", "csv", "excel", "pdf"),
                               pageLength = 10))
```


## ADT

Showing top 3 markers (by p-value) for surface protein expression.

### Heatmap of top 3 markers for each cluster

```{r, fig.width=12, fig.height=12}
Idents(s.APC) <- "wsnn_harmony_res.0.3"

markers$adt %>% 
  filter(p_val_adj < 0.0001, avg_log2FC > 0.25) %>% 
  arrange(cluster, p_val_adj, desc(avg_log2FC)) %>%
  group_by(cluster) %>% slice_head(n=3) %>%
 pull(gene) %>% unique() %>%
DoHeatmap(subset(s.APC, downsample = 1000), assay = "ADT", features = ., size = 3, angle = 90) + 
  NoLegend() +
  theme(axis.text.y = element_text(size=10))
```

### Table of top 3 markers for each cluster

```{r}
markers$adt %>% 
  filter(p_val_adj < 0.0001, avg_log2FC > 0.25) %>% 
  arrange(cluster, p_val_adj, desc(avg_log2FC)) %>%
  group_by(cluster) %>% slice_head(n=9) %>%
  DT::datatable(extensions = "Buttons",
                options = list(dom = "Blfrtip",
                               buttons = c("copy", "csv", "excel", "pdf"),
                               pageLength = 17))
```
