---
title: "integration and cell annotation"
author: "ENPRC Genomics Core"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(Seurat)
library(tidyverse)
library(here)
library(gencoreSC)
# library(RColorBrewer)
library(clustree)
library(ggpubr)
library(harmony)
knitr::opts_chunk$set(echo=FALSE, warning=TRUE, message=TRUE,
                      fig.width = 8, fig.height = 6,
                      cache=FALSE, cache.lazy = FALSE,
                      fig.align = 'center')#, include=FALSE)
```

# Report metadata

```{r}
## Report metadata
report_author <- "Micah Fletcher"
project_name <- "p21242_Satish_UM5"
report_number <- "20-APC"
report_title_root <- "subset+recluster"
report_rds_dir <- paste0("saved_rds/report-", report_number)

## Save rds here:
dir.create(here(report_rds_dir))
```

#  Load full dataset
```{r}
s.full <- read_rds(here("saved_rds/report-03/s.full-post_wnn-20230929.rds"))
sample_metadata <- read_csv(here("saved_rds/report-02/sample_metadata.csv"))
```

# Review of cell type labels

## WNN label table
```{r, fig.width=12, fig.height=12}
adts_targeted_cat_nums <- c(344651, 301071, 302363, 305651, 310951, 331231, 
                            302963, 302853, 307663, 301359, 371521, 301859, 
                            302065, 306045, 331547, 320837, 343127) %>% as.character()

adt_apriori_markers_long <- c(
  "CD4" =	"CD4 T cells, some DCs",
  "CD8a" =	"CD8 T cells, NK cells (only in NHPs)",
  "CD28" = "co-stimulatory molecule, expressed by naïve T cells and central memory T cells",
  "CD314" = "NKG2D - expressed by NK cells and some CD8 T cells",
  "CD95" = "memory T cells",
  "CD69" = "activated T cells, tissue resident T and B cells",
  "CD20" = "B cells",
  "HLA-DR" = "B cells, APCs (pDC, CD11c+ DC, CD1+ DC, CD14+ monocytes, CD16+ monocytes)",
  "CD14" = "classical and intermediate monocytes",
  "CD16" = "intermediate monocytes, non-classical/inflammatory monocytes, NK cells",
  "CD11b" = "macrophages, tissue monocytes, other innate cells",
  "CD123" = "pDC, some T cells",
  "CD1c" = "some innate cells/APCs",
  "CD11c" = "myeloid/conventional DC",
  "TCRgd" = "T cell gd",
  "CD27" = "naïve and activated effector T cells as well as NK cells and activated B cells",
  "CD7" =  "thymocytes and mature T cells")

adt_apriori_markers_short <- c(
  "CD4" =	"T CD4 + some DCs",
  "CD8a" =	"T CD8 + NK",
  "CD28" = "T naïve + T central memory",
  "CD314" = "NK + some T CD8",
  "CD95" = "T memory",
  "CD69" = "T activated + tiss res T and B",
  "CD20" = "B",
  "HLA-DR" = "B + APCs (pDC+DC+mono)",
  "CD14" = "mono classical + mono intermediate",
  "CD16" = "mono intermed + mono nonclass/inflamm + NK",
  "CD11b" = "macro + mono tiss res + other innate",
  "CD123" = "pDC + some T",
  "CD1c" = "some innate/APCs",
  "CD11c" = "myeloid/cDC",
  "TCRgd" = "Tgd",
  "CD27" = "T naïve + T act eff + NK + B act",
  "CD7" =  "thymocytes + T mature")

adt_short_name_tbl <- tibble(short_name = names(adt_apriori_markers_long), 
                   marker_of_long = adt_apriori_markers_long,
                   marker_of_short = adt_apriori_markers_short) %>%
  full_join(tibble("original_name" = rownames(s.full@assays$ADT)) %>% 
              mutate(cat_num = str_extract(original_name, "[0-9]*$"),
                     short_name = str_extract(original_name, "^[A-Z0-9a-z]*"),
                     short_name = ifelse(short_name == "TCR", "TCRgd", ifelse(short_name == "HLA", "HLA-DR", short_name)))) %>%
  mutate(marker_of_short = str_replace_all(marker_of_short, " ", "_"))
```

## WNN UMAP and dotplot

### Low-level labels
```{r, fig.width = 28, fig.height = 12}
gridExtra::grid.arrange(
  DimPlot(s.full, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.1",
          label = T, label.box = T, label.size = 6, repel = T) + 
    NoLegend() + theme(aspect.ratio=1) + ggtitle("Unsupervised clusters"),
  DotPlot(s.full, features = adt_short_name_tbl$original_name, group.by = "wsnn_harmony_res.1", cluster.idents = T) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    scale_x_discrete(labels = paste0(adt_short_name_tbl$marker_of_short, " [", adt_short_name_tbl$short_name, "]")),
  DimPlot(s.full, reduction = "wnn.umap.harmony", group.by = "ADT_and_BP_labels.wnn.res1.low", 
        label = T, label.box = T, label.size = 4, repel = T) + 
    NoLegend() + theme(aspect.ratio = 1),
  widths = c(1.5,1, 1.5)
  )
```

### High-level labels
```{r, fig.width = 28, fig.height = 12}
gridExtra::grid.arrange(
  DimPlot(s.full, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.1",
          label = T, label.box = T, label.size = 6, repel = T) + 
    NoLegend() + theme(aspect.ratio=1) + ggtitle("Unsupervised clusters"),
  DotPlot(s.full, features = adt_short_name_tbl$original_name, group.by = "wsnn_harmony_res.1", cluster.idents = T) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    scale_x_discrete(labels = paste0(adt_short_name_tbl$marker_of_short, " [", adt_short_name_tbl$short_name, "]")),
  DimPlot(s.full, reduction = "wnn.umap.harmony", group.by = "ADT_and_BP_labels.wnn.res1.high", 
        label = T, label.box = T, label.size = 4, repel = T) + 
    NoLegend() + theme(aspect.ratio = 1),
  widths = c(1.5,1, 1.5)
  )
```

# Subset B cells

```{r}
## Add metadata column for visualization
s.full@meta.data <- s.full@meta.data %>%
  mutate(APC_subset = ifelse(str_detect(ADT_and_BP_labels.wnn.res1.high, "^(APC|pDC|cDC|mono)"), 
                "keep", "remove"))

## Subset
s.APC <- subset(s.full, cells=colnames(s.full)[str_detect(s.full$ADT_and_BP_labels.wnn.res1.high, "^(APC|pDC|cDC|mono)")])
```

```{r}
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "ADT_and_BP_labels.wnn.res1.high", 
        label = T, label.box = T, label.size = 4, repel = T) + 
    NoLegend() + theme(aspect.ratio = 1)
```

# Gene expression

## Integration

For each, compare umaps grouped by animalID, group, challenge, fine ADT labels 

### Merge

```{r}
DefaultAssay(s.APC) <- "RNA"
s.APC <- NormalizeData(s.APC, verbose = F)
s.APC <- FindVariableFeatures(s.APC, selection.method = "vst", nfeatures = 2000, verbose = F)
s.APC <- ScaleData(s.APC, vars.to.regress = "CC.Difference", verbose = T)
s.APC <- RunPCA(s.APC, npcs = 30, reduction.name = "pca_rna", verbose=F)
s.APC <- RunUMAP(s.APC, dims = 1:30, reduction = "pca_rna", reduction.name = "umap_rna", verbose = F)
```

```{r}
DimPlot(s.APC, reduction = "umap_rna", group.by = "animalID") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_rna", group.by = "group") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_rna", group.by = "challenge") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_rna", group.by = "Phase") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_rna", group.by = "ADT_and_BP_labels.wnn.res1.low") + theme(aspect.ratio = 1)
```

```{r}
DimPlot(s.APC, reduction = "umap_rna", group.by = "BP.main.pruned.labels")
```

```{r, fig.width = 10, fig.height = 10}
DimPlot(s.APC, reduction = "umap_rna", group.by = "BP.main.pruned.labels", split.by = "animalID", ncol = 5)
```

### Harmony

```{r}
DefaultAssay(s.APC) <- "RNA"
s.APC <- RunHarmony(s.APC, group.by.vars = "animalID", reduction = "pca_rna", reduction.save = "harmony_rna", verbose = F)
s.APC <- RunUMAP(s.APC, reduction="harmony_rna", reduction.name = "umap_harmony_rna", dims=1:30, verbose = F)
```

```{r}
DimPlot(s.APC, reduction = "umap_harmony_rna", group.by = "animalID") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_harmony_rna", group.by = "group") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_harmony_rna", group.by = "challenge") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_harmony_rna", group.by = "Phase") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_harmony_rna", group.by = "ADT_and_BP_labels.wnn.res1.low") + theme(aspect.ratio = 1)
```

```{r}
DimPlot(s.APC, reduction = "umap_harmony_rna", group.by = "BP.main.pruned.labels")
```

```{r, fig.width = 10, fig.height = 10}
DimPlot(s.APC, reduction = "umap_harmony_rna", group.by = "BP.main.pruned.labels", split.by = "animalID", ncol = 5)
```

# Antibodies

## Integration

### Merged data

```{r}
DefaultAssay(s.APC) <- "ADT"
prots <- rownames(s.APC[["ADT"]])
VariableFeatures(s.APC) <- rownames(s.APC[["ADT"]])
# Run UMAP on adt features (not pca)
s.APC <- NormalizeData(s.APC, assay = "ADT", normalization.method = "CLR", margin = 2, verbose = F)
s.APC <- RunUMAP(s.APC, features = prots, reduction.name = 'umap_adt_feat', verbose = F)
# Run UMAP on adt pca (required to run harmony)
s.APC <- ScaleData(s.APC, verbose = F)
s.APC <- RunPCA(s.APC, features = prots, npcs = length(prots), reduction.name = "pca_adt", verbose = F)
s.APC <- RunUMAP(s.APC, reduction = "pca_adt", dims = 1:length(prots)-1, reduction.name = 'umap_adt_pca', verbose = F)
```


```{r}
DimPlot(s.APC, reduction = "umap_adt_feat", group.by = "animalID") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_adt_feat", group.by = "group") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_adt_feat", group.by = "challenge") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_adt_feat", group.by = "Phase") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_adt_feat", group.by = "ADT_and_BP_labels.wnn.res1.low") + theme(aspect.ratio = 1)
```

```{r}
DimPlot(s.APC, reduction = "umap_adt_pca", group.by = "animalID") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_adt_pca", group.by = "group") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_adt_pca", group.by = "challenge") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_adt_pca", group.by = "Phase") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_adt_pca", group.by = "ADT_and_BP_labels.wnn.res1.low") + theme(aspect.ratio = 1)
```

### Harmony

```{r}
s.APC <- RunHarmony(s.APC, group.by.vars = "animalID", reduction = "pca_adt", reduction.save = "harmony_adt", verbose = F)
s.APC <- RunUMAP(s.APC, reduction = "harmony_adt", reduction.name = "umap_harmony_adt", dims = 1:length(prots)-1, verbose = F)
```

```{r}
DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "animalID") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "group") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "challenge") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "Phase") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "ADT_and_BP_labels.wnn.res1.low") + theme(aspect.ratio = 1)
```

## Find clusters for Harmonized ADT
```{r}
s.APC <- FindNeighbors(s.APC, reduction = "harmony_adt", graph.name = c("adt.harmony_nn", "adt.harmony_snn"))
s.APC <- FindClusters(s.APC, graph.name = "adt.harmony_snn", algorithm = 1, resolution = seq(0, 1.0, 0.1))
# s.APC <- FindClusters(s.APC, graph.name = "adt.harmony_snn", algorithm = 1, resolution = seq(1.1, 1.4, 0.1))
```

```{r}
clustree::clustree(s.APC, prefix = "adt.harmony_snn_res.")
```


```{r}
s.APC@meta.data <- s.APC@meta.data %>%
  mutate(across(starts_with("adt.harmony_snn_res"), ~  factor(.x, levels(.x) %>% as.numeric() %>% sort() %>% as.character())))

DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "adt.harmony_snn_res.0.1", label = T)
DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "adt.harmony_snn_res.0.2", label = T)
DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "adt.harmony_snn_res.0.3", label = T)
DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "adt.harmony_snn_res.0.4", label = T)
DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "adt.harmony_snn_res.0.5", label = T)
DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "adt.harmony_snn_res.0.6", label = T)
DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "adt.harmony_snn_res.0.7", label = T)
DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "adt.harmony_snn_res.0.8", label = T)
DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "adt.harmony_snn_res.0.9", label = T)
DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "adt.harmony_snn_res.1", label = T)
# DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "adt.harmony_snn_res.1.1", label = T)
# DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "adt.harmony_snn_res.1.2", label = T)
# DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "adt.harmony_snn_res.1.3", label = T)
# DimPlot(s.APC, reduction = "umap_harmony_adt", group.by = "adt.harmony_snn_res.1.4", label = T)
```


# Weighted-nearest neighbors integration (GEX + ADT)

```{r}
s.APC <- FindMultiModalNeighbors(s.APC, reduction.list = list("harmony_rna", "harmony_adt"),
                                  dims.list = list(1:30, 1:16),
                                  knn.graph.name = "wknn_harmony",
                                  snn.graph.name = "wsnn_harmony",
                                  weighted.nn.name = "weighted.nn_harmony",
                                  modality.weight.name = list("mmn.weight.rna_harmony", "mmn.weight.adt_harmony"))

s.APC <- RunUMAP(s.APC, nn.name = "weighted.nn_harmony", reduction.name = "wnn.umap.harmony", reduction.key = "wnnUMAP.harmony_")
```

```{r}
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "animalID") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "group") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "challenge") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "Phase") + theme(aspect.ratio = 1)
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "ADT_and_BP_labels.wnn.res1.low") + theme(aspect.ratio = 1)
```

## Find clusters for WNN
```{r}
#s.APC <- FindClusters(s.APC, graph.name = "wsnn_harmony", algorithm = 3, resolution = seq(0.2, 2, 0.2), verbose = FALSE)
s.APC <- FindClusters(s.APC, graph.name = "wsnn_harmony", algorithm = 3, resolution = seq(0, 1.0, 0.1))
# s.APC <- FindClusters(s.APC, graph.name = "wsnn_harmony", algorithm = 3, resolution = seq(1.1, 1.4, 0.1))
```

```{r}
clustree::clustree(s.APC, prefix = "wsnn_harmony_res.")
```


```{r}
s.APC@meta.data <- s.APC@meta.data %>%
  mutate(across(starts_with("wsnn_harmony_res"), ~  factor(.x, levels(.x) %>% as.numeric() %>% sort() %>% as.character())))

DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.0.1", label = T)
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.0.2", label = T)
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.0.3", label = T)
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.0.4", label = T)
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.0.5", label = T)
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.0.6", label = T)
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.0.7", label = T)
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.0.8", label = T)
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.0.9", label = T)
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.1", label = T)
# DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.1.1", label = T)
# DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.1.2", label = T)
# DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.1.3", label = T)
# DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.1.4", label = T)
```



Going with 0.3
```{r}
Idents(s.APC) <- "wsnn_harmony_res.0.3"
```

# Cell type annotation

## WNN-based clusters

### SingleR and BluePrint database

Overview of previous report + per-label heatmaps?

### Manual labeling via ADT data

```{r, fig.weight=9, fig.height=9}
DimPlot(s.APC, reduction = "wnn.umap.harmony", group.by = "wsnn_harmony_res.0.3", label = T) + theme(aspect.ratio = 1)
```


```{r, fig.height = 8, fig.width = 7}
DotPlot(s.APC, features = adt_short_name_tbl$original_name, group.by = "wsnn_harmony_res.0.3", cluster.idents = T) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_x_discrete(labels = adt_short_name_tbl$short_name)
```

```{r, fig.height = 11, fig.width = 7}
DotPlot(s.APC, features = adt_short_name_tbl$original_name, group.by = "wsnn_harmony_res.0.3", cluster.idents = T) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_x_discrete(labels = adt_short_name_tbl$marker_of_short)
```

# Find markers

```{r}
markers <- list()
```

## RNA

```{r}
DefaultAssay(s.APC) <- "RNA"
s.APC <- s.APC %>% ScaleData(features = rownames(s.APC))
markers$rna <- s.APC %>%
  FindAllMarkers(assay = "RNA", group.by = "wsnn_harmony_res.0.3", use.test = "wilcox", 
                 logfc.threshold = -Inf, return.thresh = 1, min.pct = 0.1)
```

```{r, fig.width = 12, fig.height = 25}
markers$rna %>% 
  filter(p_val_adj < 0.0001, avg_log2FC > 0.25) %>% 
  arrange(cluster, p_val_adj, desc(avg_log2FC)) %>%
  group_by(cluster) %>% slice_head(n=10) %>% pull(gene) %>% unique() %>%
DoHeatmap(subset(s.APC, downsample=500), group.by = "wsnn_harmony_res.0.3", assay = "RNA", features = ., size = 4) + 
  theme(axis.text.y = element_text(size=8))
```

## ADT

```{r}
DefaultAssay(s.APC) <- "ADT"

markers$adt <- s.APC %>%
  FindAllMarkers(assay = "ADT", group.by = "wsnn_harmony_res.0.3", use.test = "wilcox", 
                 logfc.threshold = -Inf, return.thresh = 1, min.pct = 0.1)
```

```{r, fig.width = 12, fig.height = 25}
markers$adt %>% 
  group_by(cluster) %>%
  arrange(cluster, p_val_adj, desc(avg_log2FC)) %>%
  filter(p_val_adj < 0.0001, avg_log2FC > 0.25) %>% slice_head(n=15) %>% pull(gene) %>% unique() %>%
DoHeatmap(subset(s.APC, downsample = 500), group.by = "wsnn_harmony_res.0.3", assay = "ADT", features = ., size = 4) + 
  theme(axis.text.y = element_text(size=8))
```


# Write rds for downstream analyses

```{r}
write_rds(s.APC, here(paste0("saved_rds/report-", report_number), "s.APC.rds"))
```

# Render report

## Set filename

```{r}
report_filename <- paste0(project_name, "-", 
                          report_number, "-",
                          report_title_root, "-",
                          format(Sys.Date(), "%Y%m%d"), ".html")
```

## Render

```{r}
rmarkdown::render("20-APC-subset+recluster-format.Rmd",
                  output_file = report_filename,
                  output_dir = here("reports/"))
```



