---
title: "MAST DGE"
author: "ENPRC Genomics Core"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(Seurat)
library(tidyverse)
library(here)
library(gencoreSC)
library(DT)
library(MAST)
library(dplyr)
# library(DESeq2)
# library(gencoreBulk)

```

# Functions

```{r}
vln_plot_facet_individual <- function(obj,
                         genes,
                         assay = 'RNA',
                         subset = NULL,
                         subset_var = 'seurat_clusters',
                         grouping_var,
                         groups,
                         filter_zeros = TRUE){
  if (!is.null(subset)){
    obj <- obj[,obj@meta.data[[subset_var]] %in% subset]
  }
  mat_to_plot <- reshape2::melt(as.matrix(obj@assays[[assay]]@data))
  mat_to_plot <- mat_to_plot %>% filter(.data[['Var1']] %in% genes)
  mat_to_plot <- merge(mat_to_plot,
                       obj@meta.data %>%
                         as.data.frame() %>%
                         dplyr::select(.data[[grouping_var]], sample),
                       by.x = 'Var2', by.y = 0)
  mat_to_plot[[grouping_var]] <- factor(mat_to_plot[[grouping_var]],
                                        levels = groups)
  mat_to_plot <- mat_to_plot %>%
    filter(.data[[grouping_var]] %in% groups)

  if (filter_zeros) {mat_to_plot <- mat_to_plot %>% filter(.data[['value']] != 0)}

  ggplot2::ggplot(mat_to_plot,
                  aes(x = .data[[grouping_var]],
                      y = .data[['value']],
                      group = .data[[grouping_var]],
                      color = .data[['sample']])) +
    # geom_jitter() +
    ggplot2::geom_violin(draw_quantiles = 0.5) +
    ggforce::geom_sina(size = 0.5, alpha = 1) +
    ggplot2::facet_wrap(~`sample`)+#, scale='free_y') +
    ggplot2::theme_bw() +
    ggplot2::labs(caption = paste0('Showing expression of ', genes, 
                                   if(!is.null(subset)) {paste0('Showing expression of ', subset, ' cells\n')},
                                   if(!is.null(filter_zeros)) {'Only showing cells with non-zero expression'}),
                  y = 'Expression')
}
```

```{r cell_prop_functions}
calc_pct_expressed <- function(obj = obj.sce, features, cells, thresh.min = 0){
  round(
    x = rowSums(x = assays(obj[features, cells])$counts > thresh.min) /
      length(x = cells),
    digits = 3
  )
}

find_cells_in_group <- function(obj = obj.sce, celltype, condition = NULL, timepoint = NULL){
  obj.metadata <- as.data.frame(colData(obj))
  obj.sub <- obj.metadata[obj.metadata$cluster_labels == celltype, ]
  if (!is.null(condition)){
    obj.sub <- obj.sub[obj.sub$condition == condition,]
  }
  if (!is.null(timepoint)){
    obj.sub <- obj.sub[obj.sub$timepoint == timepoint,]
  }
  return(rownames(obj.sub))
}
```

# Load Data

```{r}
obj <- readRDS(here('saved_rds/obj_post_clustering.Rds'))
```

## Subset data?

You may only want to run DGE on select populations

```{r}
obj <- subset(obj, subset = cluster_labels == '<Cluster or label to drop>', invert = TRUE)
obj@meta.data <- droplevels(obj@meta.data)
```

# MAST Fit

I set this to check for an existing fit object first
so you don't accidentally rerun it if not needed.
This can take a while if you have many cells or celltypes

```{r}
mast_fit_filepath <- here('saved_rds/mast-results.rds')

if (file.exists(mast_fit_filepath)) {
  print('Existing RDS object found at specified path, not rerunning model fit')
  zlm_fits <- readRDS(here('saved_rds/mast-results.rds'))
} else {
  print('No saved RDS found at specified path, running model fit')
  zlm_fits <- list()
  # modify this if your populations labels are in a different metadata col
  for (celltype in levels(obj$cluster_labels)){
    # Create a subsetted object for each cell population of interest
    obj.sce_sub <- subset(obj, cluster_labels == celltype)
    obj.sce_sub@meta.data <- droplevels(obj.sce_sub@meta.data)
    obj.sce_sub <- gencoreSC::prepare_MAST_obj(obj.sce_sub, min.cells = 150)
    obj.sce_sub <- SceToSingleCellAssay(obj.sce_sub)
    
    # Define your model here. Note that you should include the CDR term 
    # (which is calculated in `gencore::prepare_MAST_obj()`)
    # per the manuscript for MAST, as this accounts for 
    # lots of technical and biological variability
    zlm_fits[[celltype]] <- zlm(
      ~ cdr +
        condition +
        timepoint +
        condition:timepoint +
        (1|capID),
      obj.sce_sub,
      method = 'glmer',
      ebayes = FALSE,
      exprs_values = 'logcounts',
      fitArgsD = list(nAGQ=0),
      parallel = TRUE,
      silent = TRUE)
  }
  saveRDS(zlm_fits, here('saved_rds/mast-results.rds'))
}

```

# Specify contrasts

As I (DG) understand how MAST calculates LogFC, I think the correct way to
specify contrasts is with the expanded terms for a given contrast. It might
not be necessary, but it also cleans up any confusion about what terms are being
included and cancelled out. There's also some weird stuff done internally for their
functions where they add the intercept to some parts and not others, so it gets
confusing and I'd prefer to just make everything explicit and let them cancel out
numerically if they should.

Anecdotally, these values made more sense too when
looking at visuals like violin plots, compared to values from LFC calculations
with simplified formulas. It is possible I simply made errors in my specifications 
or interpretations, but for now this has been the most consistent/seemingly
correct way. 

For example, even if both the numerator and denominator feature the same term 
and it should cancel out, you should include the term in both the num and denom. 
You can see the example below. 

```{r split_models_hypothesis_expansion}
contrast_numerators <- list(
  ## 1D3, post vs pre
  '`(Intercept)`+`condition1D3`+`timepointPost`+`condition1D3:timepointPost`',
  ## control, post vs pre
  '`(Intercept)`+`timepointPost`',
  ## post, 1d3 vs cnt
  '`(Intercept)`+`timepointPost`+`condition1D3`+`condition1D3:timepointPost`',
  ## pre, 1d3 vs cnt
  '`(Intercept)`+`condition1D3`',
  ## delta of delta
  '`timepointPost`+`condition1D3:timepointPost`')

contrast_denominators <- list(
  ## 1D3, post vs pre
  '`(Intercept)`+`condition1D3`',
  ## control, post vs pre
  '`(Intercept)`',
  ## post, 1d3 vs cnt
  '`(Intercept)`+`timepointPost`',
  ## pre, 1d3 vs cnt
  '`(Intercept)`',
  ## delta of delta
  '`timepointPost`')

## Give them some tidy names for extracting from lists later
contrast_names <- c(
  '1D3.post_vs_pre',
  'cnt.post_vs_pre',
  'post.1D3_vs_cnt',
  'pre.1D3_vs_cnt',
  '1d3_vs_cnt'
)
names(contrast_numerators) <- contrast_names
names(contrast_denominators) <- contrast_names


```

# Extract results

Extract results for each hypothesis and celltype of interest

```{r}
mast_results <- list()
for (celltype in names(zlm_fits)) {
  for (hyp in names(contrast_names)) {
    res <- gencoreSC::mastWaldTest(hypothesis = contrast_numerators[hyp],
                                   baseline = contrast_denominators[hyp],
                                   zlmFit = zlm_fits[[celltype]])
    test_name <- paste0(celltype, '.', hyp)
    mast_results[[test_name]] <- res
  }
}
```

## Checking summaries

Note that MAST LFC values are in the same base for the log transformation of the
data. If you used Seurat NormalizeData, it's probably natural-log transformed,
so the base is e 

```{r}
for (result in names(mast_results)){
  print(result)
  lfc = 0.25 ## log(2^.25) 
  print(paste0(nrow(mast_results[[result]] %>% filter(padj < 0.05, logFC > lfc)), ' genes significantly upregulated with lfc > ', lfc))
  print(paste0(nrow(mast_results[[result]] %>% filter(padj < 0.05, logFC < -lfc)), ' genes significantly downregulated with lfc < -', lfc))
}
# lapply(mast_results, summarizeMastResults, lfc = 0)
```

## Calculate percent expression

MAST doesn't add % expressed to outputs, but it is useful to have. This
function crudely loops through and adds % expressed for comparisons. 
It isn't very generalized right now, so you will have to go through manually
and specify what groups are in your comparisons.

```{r calc_cell_prop}
for (result_name in names(mast_results)){
  name_components <- strsplit(result_name, '.', fixed = TRUE)
  celltype <- condition <- timepoint <- NULL
  pct1 <- pct2 <- pct3 <- pct4 <- NULL
  celltype <- name_components[[1]][1]
  if (name_components[[1]][2] == 'cnt') {
    condition <- 'Control'
  } else if (name_components[[1]][2] == '1D3') {
    condition <- '1D3'
  } else if (name_components[[1]][2] == 'pre') {
    timepoint <- 'Pre'
  } else if (name_components[[1]][2] == 'post') {
    timepoint <- 'Post'
  } else if (name_components[[1]][2] == '1d3_vs_cnt') {
    cells1 <- find_cells_in_group(celltype = celltype, condition = '1D3', timepoint = 'Post') 
    cells2 <- find_cells_in_group(celltype = celltype, condition = '1D3', timepoint = 'Pre') 
    cells3 <- find_cells_in_group(celltype = celltype, condition = 'Control', timepoint = 'Post') 
    cells4 <- find_cells_in_group(celltype = celltype, condition = 'Control', timepoint = 'Pre') 
  
    pct1 <- calc_pct_expressed(features = mast_results[[result_name]]$Row.names, cells = cells1)
    pct2 <- calc_pct_expressed(features = mast_results[[result_name]]$Row.names, cells = cells2)
    pct3 <- calc_pct_expressed(features = mast_results[[result_name]]$Row.names, cells = cells3)
    pct4 <- calc_pct_expressed(features = mast_results[[result_name]]$Row.names, cells = cells4)
    
    mast_results[[result_name]] <- merge(mast_results[[result_name]], data.frame(pct1, pct2, pct3, pct4),
                                           by.x = 'Row.names', by.y = 0)
    
  } else {
    stop('portion 2 error')
  }
  if (length(name_components[[1]])>2) {
    if (name_components[[1]][3] == 'post_vs_pre'){
      cells1 <- find_cells_in_group(celltype = celltype, condition = condition, timepoint = 'Post') 
      cells2 <- find_cells_in_group(celltype = celltype, condition = condition, timepoint = 'Pre') 
    
      pct1 <- calc_pct_expressed(features = mast_results[[result_name]]$Row.names, cells = cells1)
      pct2 <- calc_pct_expressed(features = mast_results[[result_name]]$Row.names, cells = cells2)
      
      mast_results[[result_name]] <- merge(mast_results[[result_name]], data.frame(pct1, pct2),
                                           by.x = 'Row.names', by.y = 0)
  
    } else if (name_components[[1]][3] == '1D3_vs_cnt'){
      cells1 <- find_cells_in_group(celltype = celltype, condition = '1D3', timepoint = timepoint) 
      cells2 <- find_cells_in_group(celltype = celltype, condition = 'Control', timepoint = timepoint) 
    
      pct1 <- calc_pct_expressed(features = mast_results[[result_name]]$Row.names, cells = cells1)
      pct2 <- calc_pct_expressed(features = mast_results[[result_name]]$Row.names, cells = cells2)
      
      mast_results[[result_name]] <- merge(mast_results[[result_name]], data.frame(pct1, pct2),
                                           by.x = 'Row.names', by.y = 0)
    
    } else {
      stop('portion 3 error')
    }
  }
}
```

# Checking results

```{r, fig.height=6, fig.width=8}
genes <- getTopNGenesSC(mast_results[[1]],
                        min_logFC = 0.1,
                        N = 6, 
                        min_padj = 0.05,
                        pval_col = 'padj', lfc_col = 'logFC', gene_name_col = 'Row.names') 

mast_results[[1]] %>% filter(Row.names %in% genes)
```

Some helper functions from the gencoreSC package

```{r, fig.height=6, fig.width=8}
gcoreVlnPlot(obj = obj,
             genes = genes,
             assay = 'RNA', 
             subset_var = 'cluster_labels',
             subset = 'Monocytes',
             grouping_var = 'condition.timepoint',
             groups = c('Pre.1D3', 'Post.1D3', 'Pre.control', 'Post.control'))
```


```{r, fig.height=6, fig.width=8}
gcoreVlnPlot_facetted(obj,
                      genes = genes[[1]],
                      facet_var = 'sample',
                      subset_var = 'cluster_labels', 
                      subset = 'Monocytes',
                      grouping_var = 'condition.timepoint',
                      groups = c("Pre.1D3","Post.1D3","Pre.control","Post.control"))
```

# Write sheets

```{r}
writeDGEResults(mast_results,
                drop_NA = FALSE,
                outdir = here('outputs/'),
                output_name = 'dge.xlsx',
                p_val_colname = 'p_val_adj',
                drop_NA = FALSE,
                write_rownames = TRUE)
```
