---
title: "p23120 CD8 T cell DGE"
author: "ENPRC Genomics Core"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(Seurat)
library(tidyverse)
library(here)
library(gencoreSC)
library(DT)
library(dplyr)
# library(DESeq2)
# library(gencoreBulk)

```

# Load seurat obj

```{r}
obj <- readRDS(here('saved_rds/obj_post_clustering.Rds'))
```

# Label clusters for composition analysis

```{r}
cluster_identities <- c(
  'pDC', #0
  'pDC',    
  'cDC2',    
  'Monocytes',
  'other DCs',      
  'pDC',        
  'cDC1', 
  'pDC',    
  'other DCs',    
  'Activated DCs', #10
  'other DCs'
)
obj$previous_labels <- obj$cluster_labels
obj$cluster_labels <- plyr::mapvalues(
  obj$cluster_labels, 
  from = levels(obj$cluster_labels),
  to = cluster_identities
)
Idents(obj) <- obj$cluster_labels
```

# Celltype tables

These haven't been generalized. You will have to tweak them to add metadata
as needed

```{r}
celltype_count_table <- merge(as.data.frame(table(obj$capID, obj$cluster_labels)) %>%
  reshape2::dcast(Var1 ~ Var2, value.var = 'Freq'),
  project_metadata, by.x = 'Var1', by.y = 'Capture') %>%
  select(-c(Time,Cells,Libraries,ADT_panel,Notes)) %>% 
  relocate(Var1, Sample.s., ATI, Condition)
celltype_count_table
```

```{r}
celltype_count_table_2 <- celltype_count_table %>%
  select(-c(Var1, Sample.s.)) %>%
  group_by(Condition, ATI) %>%
  summarise_each(list(sum))
celltype_count_table_2
```


# Generate data

scCoda needs a table with a row for each sample and a column for each 
metadata variable and cell population you want to model

This will need to be tailored to the metadata you're interested in modeling

```{r}
comp_table <- obj@meta.data %>% 
  group_by(full_labels, condition, timepoint, sample) %>%
  summarise(n = n()) %>%
  pivot_wider(id_cols = c(sample, condition, timepoint),
              names_from = full_labels, values_from = n) %>%
  arrange(sample, timepoint) %>%
  mutate_all(~replace(., is.na(.), 0))
write.table(x = comp_table, 
            file = here('outputs/composition_table.txt'),
            quote = FALSE,
            row.names = FALSE, col.names = TRUE, sep = '\t')
```

# scCoda ouptut

See the README in 'helper_scripts' on how to run the jupyter notebook for scCODA

```{r}
sccoda_results <- read.csv(here('outputs/scCODA_output.csv'))
sccoda_results %>% 
  arrange(desc(Inclusion.probability))
```


## boxplots

```{r}
is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}
```

This needs to be generalized, just dumping it here for now as an example

```{r, fig.height=4, warning=FALSE, fig.width=15}
pre_vs_post_coarse_comp_boxplot <- full_comp_table %>%
  group_by(full_labels, timepoint) %>%
  mutate(outlier = ifelse(is_outlier(prop), sample, NA)) %>%
    mutate(full_labels = factor(full_labels)) %>%
  mutate(full_labels = relevel(full_labels, 'Other (non-DCs)')) %>%
  ggplot(aes(x = group, y = prop, group = group, color = group, shape = sample)) + 
  # ggrepel::geom_text_repel() +
  # scale_color_manual(values = c('Post' = my_color_pal$pre, 'Pre' = my_color_pal$post)) +
  geom_boxplot(outlier.colour = NA, show.legend = FALSE) +
  ggforce::geom_sina(size = 3) +
  labs(x = element_blank(), y = 'Proportion of sample', color = 'Group',
       caption = 'Portions are of the total capture for an individual.\n* = significant below 0.05 FDR') +
  theme_bw() +
  scale_shape_manual(values = c(1:5))+ 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  facet_grid(~ factor(full_labels)) +
  ## This is known to be significant from the scCODA results
  geom_signif(data = data.frame(full_labels = 'pDC',
                                label = '*',
                                y = .035,
                                start = 1.5, end = 3.5,
                                outlier = NA),
              aes(annotations = label, y_position = y,
                  xmin = start, xmax = end),
              inherit.aes = FALSE,
              tip_length = -0.03,
              margin_top = -0.02,
              vjust = 2.5,
              manual = TRUE, map_signif_level = TRUE)
pre_vs_post_coarse_comp_boxplot
ggsave(here('outputs/innate/DC+monocyte/composition_boxplot.png'), height = 4, width = 15)
```

